Thesis - Upwelling and MHW/Cold spells analyses

### 2019-08-28 Write_up/write_up.Rmd
wind_analyses
fixed

With global warming continuing at a rapid pace, the amount of greenhouse gases within the atmosphere will subsequently increase in the future. This increase will undoubtedly have an impact on the world’s oceans and the marine life within it,  both directly and indirectly. The oceanic process of coastal upwelling is one aspect that is likely to be affected as changes in global temperature will affect temperature and pressure gradients, and in turn affect alongshore winds which are responsible for driving the upwelling process within EBUS regions. Because upwelling is important for productivity within coastal systems, a decrease in this process will negatively impact marine fauna and flora as well as human activities such as fishing. To assess changes in upwelling events, the aim of this investigation is to use SST, wind, and chlorophyll data to monitor the durations and intensities of upwelling events within the Benguela EBUS and to identify whether this resulted in an extreme anomalous cold event. Atmospheric-ocean coupling is known to be affected by climate change, likely changing wind patterns and leading to upwelling favourable conditions (García-Reyes et al., 2015). Here, it is tested whether or not we see an increase in the frequency and intensity with EBUS. We also aim to determine whether or not there is a change in the rate of onset of marine heatwaves and marine coldspells.

@@

The primary aim of the first study was to make use of various remotely sensed SST data, in-situ collected coastal seawater temperature data and chlorophyll-a data. It was also useful to examine how the intensity and duration of upwelling varies per dataset as these datasets are collected at different resolutions, with the assumption that high resolution data may reveal a more accurate detection of the duration and intensity of upwelling events. Wind data used in these analyses were predicted to show clear patterns between weak and favourable upwelling winds and its impact on the intensity and duration of upwelling. Chlorophyll a data were then used to examine whether or not the intensity and duration of upwelling may lead to an increase or a decrease in the chlorophyll a concentration at the various sites. It was hypothesised that an increase in the intensity and duration of upwelling will show an increase in the abundance of chlorophyll a concentration. The testing of the hypothesis would be made possible with a series of analyses that allows one to select specific time periods from a range of datasets. Thereafter analyses rate of onset of MHW and MCS was done.

Thereafter the aim of the second study was to determine the occurrence of MCSs along the west coast of Southern Africa between the nearshore (<400m from the shore) and offshore. The most appropriate dataset for the measurement of the nearshore temperatures within South Africa was the South African Coastal Temperature Network (SACTN) dataset (Smit et al. 2013). With the knowledge obtained from study 1, it was now possible to investigate whether a more intense upwelling event may lead to an increase in the intensity and duration of MCSs. The SACTN dataset and a total of five SST datasets could be used to investigate the occurrence of marine cold-spells (MCSs) and marine heatwaves with the methodology introduced by Hobday et al. 2016. This research could also assists in the understanding of local oceanographic patterns along the west coast of South Africa and the impact anthropogenically forced climate change could have for the recent changes in patterns and long-term trends.

@@

The primary aim of the second study was to make use of remotely sensed satellite sea surface temperature for the four EBUS to determine wether or not a change exsit in the rate of onset of marine coldspells and marine heatwaves. It was also useful to examine how the intensity and frequency of upwelling events varied between datasets and years given the influence of climate change and sea surface temperature rises.

>>>>

With global warming continuing at a rapid pace, the amount of greenhouse gases within the atmosphere will subsequently increase in the future. This increase will undoubtedly have an impact on the world’s oceans and the marine life within it, both directly and indirectly. The oceanic process of coastal upwelling is one aspect that is likely to be affected as changes in global temperature will affect temperature and pressure gradients, and in turn affect alongshore winds which are responsible for driving the upwelling process within EBUS regions. Because upwelling is important for productivity within coastal systems, a decrease in this process will negatively impact marine fauna and flora as well as human activities such as fishing. To assess changes in upwelling events, the aim of this investigation is to use SST and wind data to monitor the durations and intensities of upwelling events within the Benguela EBUS and to identify whether this resulted in an extreme anomalous cold event. Atmospheric-ocean coupling is known to be affected by climate change, likely changing wind patterns and leading to upwelling favourable conditions (García-Reyes et al., 2015). Here, it is tested whether or not we see an increase in the frequency and intensity with EBUS. We also aim to determine whether or not there is a change in the rate of onset of upwelling and marine coldspells.

@@

The primary aim of the first study was to make use of various remotely sensed SST and in-situ collected coastal seawater temperature data. The most appropriate dataset for the measurement of the nearshore temperatures within South Africa was the South African Coastal Temperature Network (SACTN) dataset (Smit et al. 2013). It was useful to examine whether or not the same upwelling patterns exist at different distances from the coastline and whether or not the same patterns exist between the different datasets.

@@

The primary aim of the second study was to make use of remotely sensed satellite sea surface temperature for the four EBUS to determine whether or not a change exsit in the rate of onset of marine coldspells and marine heatwaves. It was also useful to examine how the intensity and frequency of upwelling events varied between datasets and years given the influence of climate change and sea surface temperature rises. Wind data used in these analyses were predicted to show clear patterns between weak and favourable upwelling winds and its impact on the intensity and duration of upwelling. The testing of the hypothesis would be made possible with a series of analyses that allows one to select specific time periods from a range of datasets. Thereafter analyses rate of onset of MCS was done.This research could also assists in the understanding of local oceanographic patterns along the west coast of South Africa and the impact anthropogenically forced climate change could have for the recent changes in patterns and long-term trends
